Allocation Removal by Partial Evaluation in a Tracing JIT

### 2010-10-12 talk/pepm2011/paper.tex
added refs
unify stuff
Write a small bit of text for "Analysis of the Algorithm".

XXX algorithm is linear in the length of the trace

XXX Category 2 The optimization of
Section~\ref{sec:statics} deals with them too: the \texttt{new} that creates them and
the field accesses are deferred, until the point where the object escapes.

XXX optimization particularly effective for chains of operations

@@

Prolog

@@

>>>>

While we do not offer a formal proof of it, it should be relatively clear
that the algorithm presented above is sound: it works by delaying (and
often completely removing) some operations.  The algorithm runs in a
single pass over the list of operations.  We can check that altough
recursively lifting a static object is not a constant-time operation,
the algorithm only takes a total time linear in the length of the trace.
Moreover, it gives the ``best'' possible result within its constrains,
e.g.\ in term of the number of residual operations.  The
algorithm itself is not particularly complex or innovative; our focus is
rather that \emph{in the context of tracing JITs} it is possible to find a
simple enough algorithm that still gives the best results.

Note in particular that objects in category 1 (i.e.\ the ones that do
not escape) are completely removed; moreover, objects in category 2
(i.e.\ escaping) are still partially dealt with: if such an object
escapes later than its creation point, all the operations inbetween that
involve the object are removed.

The optimization is particularly effective for chains of operations.
For example, it is typical for an interpreter to generate sequences of
writes-followed-by-reads, where one interpreted opcode writes to some
object's field and the next interpreted opcode reads it back, possibly
dispatching on the type of the object created just before.  In the case
of PyPy's Python interpreter, this optimization can even remove the
allocation of all intermediate frames that occur in the interpreter,
corresponding to all calls that have been inlined in the trace.

@@

Prolog:
partially static datastructures are already built-in to Prolog and was built-into partial
evaluation from the beginning \cite{lloyd_partial_1991}.
 FP

@@

xxx: relation to compile-time garbage collection \cite{mazur_practical_2001};  separation logic; John Hughes: type specialization
